The role of advection versus diffusion in intraventricular transport

### 2025-02-04 paper/main.tex
appendix edits, discussion on role of cilia

\documentclass[fleqn]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage{lipsum}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{siunitx}
\usepackage{physics}
\usepackage{stmaryrd} % double brackets for jump operator
\usepackage{nameref}
\usepackage{booktabs}
\usepackage{makecell} % for table cells with line breaks
\usepackage{bm}
\usepackage{amsopn}
\usepackage[skip=10pt]{caption}
\usepackage{color}
\usepackage{xcolor,colortbl}
\usepackage[nameinlink, capitalise]{cleveref}
\usepackage{graphicx}
\usepackage{graphbox} % for center-aligning figures vertically
\usepackage{tkz-euclide} % draw tikz pictures
\usepackage{subcaption}
\usepackage{tablefootnote} % for table footnotes
\usepackage{float} % figure placement

@@

In this study, we combined computational and experimental approaches to investigate the role of motile cilia in CSF flow generation and solute transport.
First, we computed flow patterns and intraventricular transport in silico. To this end, we used the geometry of the zebrafish larval brain ventricles,
for which we have detailed knowledge of cilia properties and CSF movement velocities and frequencies.
Notably, this early developmental stage allows us to simplify our model by omitting CSF secretion, which starts at later stages, while still being well conserved with mammals.
Next, we validated our computation model with experimental data by monitoring the dynamics of a photoconvertible fluorescent protein expressed within the brain ventricles.
Our results identified that diffusion plays a major role in the transport of small solutes, whereas cilia majorly contribute to the movement of larger particles.
We also found that the location of cilia and the ventricular geometry greatly impact solute distribution.
Altogether, our work presents a computational framework and new concepts of how cilia collaborate with other factors to distribute molecules and particles within the brain ventricles.

\section*{Methods}

\mer{@HH: I suggest you write \cref{fig:figure1}a, rather than \cref{fig:figure1} panel \textbf{a}, throughout. Increases signal-to-noise ratio.}

\subsection*{Image-based computational geometries of zebrafish brain ventricles}
We generated a 3D representation of the brain ventricular walls using confocal imaging data of a zebrafish embryo injected intraventricularly with a 70 kDa dye at 2 dpf~\cite{Olstad2019CiliaryDevelopment}. 
From the imaging data, we constructed a computational mesh of the ventricles (\Cref{fig:figure1}d). The interior volume was meshed with fTetWild~\cite{Hu2020FastWild}. 
The \emph{standard} mesh consisted of 132 134 tetrahedral cells and had a maximal (minimal) edge length of 14.9 $\mu$m (4.68 $\mu$m). 
At 2 dpf, the ventricular system consists mainly of 3 cavities: the telencephalic  (anterior), the di-/mesencephalic  (middle) and the rhombencephalic (posterior) ventricles, which are connected by ducts. 
Parts of the ventricular surfaces were marked to be lined with motile cilia~\cite{Olstad2019CiliaryDevelopment} (\Cref{fig:figure1}d, pink markers).
\begin{figure}%[hbt!]
    \centering
    \includegraphics[width=\linewidth]{graphics/figure1.png}
    \caption{\lyng{Experimenting with velocity colorbars, thoughts on the two options? Or have two separate as it is now?}}
    \label{fig:figure1}
\end{figure}

@@

\subsection*{Computational CSF dynamics}
The beating motion of the cilia generates CSF flow with steady rotational structures~\cite{Olstad2019CiliaryDevelopment} within the brain ventricles (\Cref{fig:figure1}a--c).
In addition, cardiac pulsations induces pulsatile CSF flow. We model this flow of CSF in the brain ventricles by the time-dependent, incompressible Stokes equations which read as follows.
Find the CSF velocity $\uu = \uu(\xx, t) = (u_x(\xx, t), u_y(\xx, t), u_z(\xx, t))$ and the CSF pressure $p = p(\xx, t)$ for $\xx \in \Omega$ and time $t>0$, such that
\begin{subequations}
    \begin{align}
      \rho \pdifft{\uu} - \nabla \cdot \bsig &= \mathbf{0}
      \quad \text{in } \Omega,
      \label{eq:stokes_eq_mom}\\
      \nabla \cdot \uu &= 0
      \quad  \text{in } \Omega,
  \end{align}
  \label{eq:stokes_eqs}% 
\end{subequations}%
where the stress tensor is defined as $\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$ and $\beps(\uu) = \frac{1}{2}\left(\nabla \uu + (\nabla\uu)^T\right)$ is the strain-rate tensor. 
Bold-face characters denote vectors or tensors, and $\mathbf{I}$ is the identity tensor in three dimensions. Furthermore, $\rho=1000 \ \mathrm{kg/m^3}$ is the CSF density and $\mu=6.97\times 10^{-4}$ Pa is the dynamic viscosity of the CSF.
We introduce the tangential traction $\bsigpar$ as the tangential component of the traction $\hat{\bsig}=\bsig\nn$:
\begin{equation*}
    \bsigpar = P_{\nn}(\hat{\bsig}) = (\mathbf{I} - \nn\otimes\nn)\hat{\bsig},
\end{equation*}
where $\nn$ is the outer unit normal of the boundary surface $\Gamma = \partial\Omega$, and $P_{\nn}(\rr)$ is the tangential  projection of a vector $\rr$ onto $\Gamma$.
We consider three types of boundary conditions:  (i) \emph{tangential traction} (no flow normal to the boundary $\uu \cdot \nn = 0$ with a force $\btau$ applied tangentially on the boundary $\bsigpar = \btau$),
(ii) \emph{normal pressure} (a pressure $\Tilde{p}$ applied in the normal direction: $(\mu\nabla\uu - p\mathbf{I})\nn = \Tilde{p}(t)\nn$),
and (iii) \emph{free slip} (no flow normal to the boundary $\uu \cdot \nn = 0$ and no tangential forces $\bsigpar = {\bf 0}$).
Unless noted otherwise, the free-slip condition is applied on the boundary. We assume that the system starts at rest ($\uu(\xx, t=0) = \mathbf{0}$).

\subsection*{Cilia-driven flow}
On the ciliated regions of the ventricle walls $\Gc$ (\Cref{fig:figure1}d), we impose a constant-in-time tangential traction $\bsigpar=\btau$ to represent the net forces of the cilia acting on the CSF.
We set $\btau(\xx, \rr) = \tau \lambda(\xx) P_{\nn}(\rr)$, where the function $\lambda(\xx)$ and the sign of the vector $\rr=\pm(1, 0, 1)$ varies depending on the cilia region (see~\Cref{subsec:appendixA1}).
Both $\rr$ and the value $\tau=6.5\times 10^{-4}$ Pa were chosen by model calibration against the experimental data~\cite{Olstad2019CiliaryDevelopment}.

\subsection*{Flow induced by cardiac pulsations}\label{subsec:pressure_bcs}
To drive a cardiac-induced pulsatile flow inside the brain ventricles, we set $\Tilde{p}(t)=0$ and $\Tilde{p}(t)=-A\sin{\omega t}$ on anterior and posterior parts of the boundary $\Gp$~(\Cref{fig:figure1}d, blue markers).
The amplitude $A=1.5\times 10^{-3}$ Pa determines the magnitude of the pulsatile flow, and was calibrated based on experimental data~\cite{Olstad2019CiliaryDevelopment}.
The angular cardiac frequency $\omega=2\pi f=6.97$ rad/s was set based on a measured cardiac frequency of $f=2.22$ Hz.

\subsection*{Computational model of solute transport after photoconversion}
We simulate transport of photoconverted proteins within the brain ventricles by modeling transport of a solute with concentration $c(\xx, t)$ for $\xx\in\Omega$ and time $t > 0$ by the advection-diffusion equation
\begin{equation}
    \pdifft{c} + \nabla\cdot\JJ = 0 \quad \text{in } \Omega,
    \label{eq:adv_diff_strong}
\end{equation}
where the total concentration flux $\mathbf{J}$ is assumed to consist of an advective and a diffusive flux:
\begin{equation*}
    \JJ = c\uu - D\nabla c .
\end{equation*}
Here, $\uu$ is the CSF velocity field, governed by the Stokes equations~\eqref{eq:stokes_eqs}, and $D$ is the molecular diffusion coefficient of the solute. Initially, we set $c(\xx, 0)=0$ for all $\xx\in\Omega$.
The ventricular walls are assumed to be impermeable (admitting \emph{no flux} i.e.~$\JJ\cdot\nn=0$), except at the anterior/posterior boundary $\Gamma_p$.
\fixme{At the anterior/posterior boundary we allow the solute concentration to be advected into and out of the domain in a manner that ensures conservation of mass (see~\Cref{sec:appendix:X} for details).}
\mer{MER: @Halvor: could you have a go at moving the precise description to the Appendix?}
\fixme{There, we consider either prescribed \emph{influx} in the form of a time-periodic condition on the total flux $(\JJ\cdot\nn)|_{\mathrm{in}}^{n+1}=(\JJ\cdot\nn)|_{\mathrm{out}}^{n}$ and/or \emph{outflux} with zero diffusive flux $D\nabla c\cdot\nn=0$.}
\fixme{Additionally, when considering pulsatile motion, the in- and outflow on the pressure boundary $\Gp$ must be accounted for. We split $\Gp$ in two parts; $\Gin$ denotes the part of $\Gp$ where $\uu\cdot\nn < 0$;
$\Gout$ denotes the part where $\uu\cdot\nn > 0$. On $\Gin$, a total flux boundary condition was set: the inflow flux at a given timestep was set equal to the outflow boundary flux at the previous timestep.
This was done to approximate a periodic boundary condition. On $\Gout$, we considered the diffusive flux normal to the boundary to be zero.}

To computationally represent protein photoconversion in a region $\Omega_{\mathrm{pc}} \subset \Omega$ (corresponding to ROI 1), we prescribe a given time-dependent value to the solute concentration in this region. 
Mimicking the experimental protocol, we consider the dorsal part of the diencephalic ventricle as the subdomain $\Omega_{\mathrm{pc}}$, and set
\begin{equation}
  c(\xx, t)=\frac{\log{(1+t/a)}}{\log{(1+T/a)}}
  \qquad \xx \in \Omega_{\mathrm{pc}}, \quad \fixme{t > 0}.
\label{eq:injection_curve}
\end{equation}
The value $a=65$ was chosen to fit the fluorescence intensity curve observed in the physical photoconversion experiments.

@@

\subsection*{Physical parameters}
The CSF density $\rho$ and dynamic viscosity $\mu$ were set following Bloomfield \textit{et al.}~\cite{Bloomfield1998EffectsFluid}~(\Cref{tab:parameter_table}). We considered large, medium and small-sized molecules by three different diffusion coefficients $D_1, D_2$ and $D_3$ (\Cref{tab:parameter_table}). Experimental data~\cite{Olstad2019CiliaryDevelopment} indicate \fixme{velocities} proximate to the ventricular walls of $27.4 \pm 5.4$ \textmu m/s and $2.6 \pm 0.6$ \textmu m/s in the dorsal and ventral regions of the diencephalic ventricle, respectively. In the telencephalic ventricle, the velocities were estimated at $4.7 \pm 1.3$ \textmu m/s. We calibrated $\tau$ and $A$ heuristically by comparing these experimental velocity values with the simulated, maximum velocity magnitude. In addition, particle tracking data of pulsatile motion was used in the calibration process to determine the relative contributions of the cardiac pulsations, as compared to the cilia-driven flow, to the velocity magnitude.
\mer{MER: What velocities? Net velocities? Velocity magnitude aka speed? Also, haven't you said this part about the experimental data already? Either place the details there, or all here. I suggest there.}
\mer{MER: I suggest you introduce parameter values when introduced, and move the parts relevant for cilia/cardiac flow to where you mentioned it first, and then merge this section with the Estimation of diffusion .... Then remove this section and \Cref{tab:parameter_table}.}
\begin{table}[h!]
    \centering
    \begin{tabular}{lcccc}
    \toprule
        Parameter & Symbol & Value & Unit & Reference \\
        \midrule
        Fluid density & $\rho$ & $1000$ & $\mathrm{kg/m^3}$ & \cite{Bloomfield1998EffectsFluid} \\
        Dynamic viscosity & $\mu$ & $\num{6.97e-04}$ & $\mathrm{Pa\cdot s}$ & \cite{Bloomfield1998EffectsFluid} \\
        Cardiac frequency & $f$ & 2.22 & Hz &  \cite{Olstad2019CiliaryDevelopment}\\
        Angular cardiac frequency & $\omega = 2\pi f$ & 6.97 & rad/s & \cite{Olstad2019CiliaryDevelopment}\\
        Cilia tangential traction & $\tau$ & \num{6.5e-04} & Pa & * \\
        Pressure amplitude & $A$ & \num{1.5e-03} & Pa & * \\
        %Ref. frame motion amplitude & $A_{\mathrm{ref}}$ & \num{9.19e-04} & m/s & *\\
        Exosome diffusion coefficient & $D_1$ & $\num{1.63e-12}$ & $\mathrm{m^2/s}$ & \textdagger\\
        STM+GFP diffusion coefficient  & $D_2$ & $\num{5.75e-11}$ & $\mathrm{m^2/s}$ & \textdagger\\
        Dendra2 diffusion coefficient  & $D_3$ & $\num{1.15e-10}$ & $\mathrm{m^2/s}$ & \textdagger\\
    \bottomrule
    \end{tabular}
    \caption{Physical parameters of the flow and transport models. (*) Calibrated based on experimental data~\cite{Olstad2019CiliaryDevelopment}. (\textdagger) Estimated based on experimental data~\cite{GuraSadovsky2017MeasurementExpansion, Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells}.}
    \label{tab:parameter_table}
\end{table}

@@

\mer{MER: @Halvor I suggest you move these two paragraphs into their respective Results sections (Results 2 and 5). Putting the bullet points above at the end of the Stokes equations is a great idea!}
When simulating transport of solute in the ventricles, we consider the same six different regions of interest that were considered in the photoconversion experiments. In region number one (the photoconversion region), we use the function~\eqref{eq:injection_curve} to enforce the concentration such that it represents the experimental data of photoconversion of proteins. We consider the three different diffusion coefficients introduced in previous sections to simulate and compare results for transport of molecules of varying sizes.

@@

\subsection*{Quantities of interest}
We calculate the mean solute concentrations
\begin{equation}
    \cbar_i(t) = \frac{1}{|\Omega_i|}\int_{\Omega_i} c(\xx, t)\,\dx,
    \label{eq:c_mean_i}
\end{equation}
in the six regions of interest (ROIs) $\Omega_i$ ($i = 1, \dots, 6$) as functions of time, where $|\Omega_i|$ is the volume of the domain $\Omega_i$. The dynamics of ROI 1 follows directly from~\eqref{eq:injection_curve}. In addition, we report time-to-threshold: the time $\hat{t}_i$ when the mean concentration $\cbar_i$ first exceed a threshold value $\hat{c}_i$. We use $\hat{c}_i=0.25$ for ROI 1--4 and $\hat{c}_i=0.10$ for ROI 5--6.

@@

Transport of solute is quantified by the profile of the mean concentrations $\cbar_i(t)$ in the regions of interest (ROIs) $\Omega_i$ as functions of time~(\Cref{fig:fig2}b--d).
An illustration of the ROIs is provided in \Cref{fig:fig2}e. The dynamics of ROI 1 follow from imposing~\Cref{eq:injection_curve}.
For ROIs 2--6, we observe more rapid concentration increases when the diffusion coefficient increases. 
For all three diffusion coefficients, the dynamics of ROI 6 is the slowest.
Region 5 exhibits a more rapid increase in mean concentration, and an even faster increase is observed for ROIs 2--4.
The profiles of the latter three regions are similar, displaying only small differences in all of the three plots.
\lyng{HH: Add some quantification in terms of percentages, e.g. final values of $\cbar_i$ for the different $D_i$?}
In region 6, the threshold is never exceeded for $D_1$, with the mean concentration close to zero at the end of the simulation. 
To quantify the balance of advection versus diffusion as transport mehcanisms on a global scale, we calculate the Péclet numbers $\mathrm{Pe}_1=519$, $\mathrm{Pe}_2=14.7$ and $\mathrm{Pe}_3=7.4$ for $D_1$, $D_2$ and $D_3$, respectively.
Since $\mathrm{Pe}>1$, this indicates that advection is the dominant transport mechanism for the molecules considered.

@@

\subsection*{Loss of cilia motility in specific subpopulation slows down transport}
We previously identified three distinct ciliated cell lineages in the
zebrafish larval brain~\cite{Olstad2019CiliaryDevelopment, DGama2025MotileBrain}.
To identify the role of these lineages on CSF transport, we modified flow boundary conditions
in the three different ciliated regions (\Cref{fig:figure5_compare_cilia_modifications}a) and simulated
CSF flow using the Stokes equations. We identified that the velocities were most strongly
affected when removing cilia in the dorsal diencephalic cilia,
as the maximum velocity dropped from 27.7 $\mu$m/s to 8.0 $\mu$m/s.
In contrast, removing telencephalic and ventral diencephalic had minimalno to little impact and resulted in maximum velocities of 27.7 $\mu$m/s or 27.2 $\mu$m/s, respectively.

@@

\subsection*{The roles of advection and diffusion in intraventricular solute transport}
To determine the roles of advection and diffusion in solute transport, we quantified transport
based on changes in fluorescence intensity $\Delta F$ (experiments) and
mean concentration $\cbar$ (simulations) as functions of time.
Even though we identified very similar dynamics in simulations and
\emph{in vivo} experiments, the dynamics of the simulations were generally
more rapid and resulted in higher total amounts.
Differences between experiments and simulations may relate to
ventricular size, as we showed that modifying the duct volumes impacts
solute distribution and the geometry used for the model was larger than in experiments.
Alternatively, it may relate to the way we defined the photoconversion site (region of interest 1)
in the simulations, since the volume of the photoconversion region determines
the total amount of solute, and thereby the rate of transport. Another source
of discrepancy may relate to the permeability of the ventricular wall,
which we have ignored in our simulation. Reports have shown that the
neuroepithelium lining the walls of brain ventricles in 24h old zebrafish
is permeable to dyes smaller than 70 kDa~\cite{Chang2012AnNeuroepithelium}.
Since Dendra2 is around 26 kDa in size~\cite{Gurskaya2006EngineeringLight},
and thus under this reported limit, one could assume that if Dendra2
leaks out of the brain ventricles, the dynamics become slower.

@@

%The error norms used to calculate the convergence rates that we report are not the ones naturally appearing in \emph{a priori} error estimates of a $\mathrm{BDM}_k$-$\mathrm{DG}_{k-1}$ discretization. Therefore, we cannot say anything explicit about expected convergence rates. However, the polynomial order $k$ of the elements provide an upper bound on the convergence rate, with the expectation that the $L^2$ error norm convergence rates should not exceed $k+1$. We observe convergence rates approaching $k+1$ and $k$ for the $\uu$ and $p$ error norms in $L^2$, when using polynomial orders $k$ and $k-1$, respectively.

>>>>

\documentclass[fleqn]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage{lipsum}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{siunitx}
\usepackage{physics}
\usepackage{stmaryrd} % double brackets for jump operator
\usepackage{nameref}
\usepackage{booktabs}
\usepackage{bm}
\usepackage{amsopn}
\usepackage[skip=10pt]{caption}
\usepackage{color}
\usepackage{xcolor,colortbl}
\usepackage[nameinlink, capitalise]{cleveref}
\usepackage{graphicx}
\usepackage{tkz-euclide} % draw tikz pictures
\usepackage{subcaption}
\usepackage{tablefootnote} % for table footnotes
\usepackage{float} % figure placement

@@

In this study, we combined computational and experimental approaches to investigate the role of motile cilia in CSF flow and solute transport.
First, we computed flow patterns and intraventricular transport \emph{in silico}. To this end, we used the geometry of the zebrafish larval brain ventricles,
for which we have detailed knowledge of cilia properties and CSF movement velocities and frequencies.
Notably, this early developmental stage allows us to simplify our model by omitting CSF secretion, which starts at later stages, while still being well-conserved with mammals.
Next, we validated our computation model with experimental data by monitoring the dynamics of a photoconvertible fluorescent protein expressed within the brain ventricles.
Our results show that diffusion plays a major role in the transport of small solutes, whereas cilia majorly contribute to the movement of larger particles.
We also found that the location of cilia, as well as the ventricular geometry, greatly impact solute distribution.
Altogether, our work presents a computational framework together with new concepts of how cilia
collaborate with other factors to distribute molecules and particles within the brain ventricles.

\section*{Methods}
\subsection*{Image-based computational geometries of zebrafish brain ventricles}
To model flow and transport in zebrafish larval brain ventricles, we generated a 3D representation of the brain ventricular walls using confocal imaging data of a zebrafish embryo
injected intraventricularly with a 70 kDa dye at 2 dpf~\cite{Olstad2019CiliaryDevelopment}. 
From the imaging data, we constructed a computational mesh of the ventricles (\Cref{fig:figure1}d).
The interior volume was meshed with fTetWild~\cite{Hu2020FastWild}. 
The \emph{standard} mesh consisted of 132 134 tetrahedral cells and had a maximal (minimal) edge length of 14.9 $\mu$m (4.68 $\mu$m). 
At 2 dpf, the ventricular system consists mainly of 3 cavities: the telencephalic (anterior),
the di-/mesencephalic (middle) and the rhombencephalic (posterior) ventricles, which are connected by ducts. 
Parts of the ventricular surfaces were marked to be lined with
motile cilia~\cite{Olstad2019CiliaryDevelopment} (\Cref{fig:figure1}d, pink markers).
\begin{figure}%[hbt!]
    \centering
    \includegraphics[width=\linewidth]{graphics/figure1.png}
    \caption{\lyng{Experimenting with velocity colorbars, thoughts on the two options? Or have two separate as it is now?}}
    \label{fig:figure1}
\end{figure}

@@

\subsection*{Computational CSF dynamics}
The beating motion of the cilia generates CSF flow with steady rotational structures~\cite{Olstad2019CiliaryDevelopment}
within the brain ventricles (\Cref{fig:figure1}a--c).
In addition, cardiac pulsations induces pulsatile CSF flow.
We model this flow of CSF in the brain ventricles by the time-dependent,
incompressible Stokes equations which read as follows.
Find the CSF velocity $\uu = \uu(\xx, t) = (u_x(\xx, t), u_y(\xx, t), u_z(\xx, t))$
and the CSF pressure $p = p(\xx, t)$ for $\xx \in \Omega$ and time $t>0$, such that
\begin{subequations}
    \begin{align}
      \rho \pdifft{\uu} - \nabla \cdot \bsig &= \mathbf{0}
      \quad \text{in } \Omega,\\
      \nabla \cdot \uu &= 0
      \quad  \text{in } \Omega,
  \end{align}
  \label{eq:stokes_eqs}% 
\end{subequations}%
where the stress tensor is defined as $\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$
and $\beps(\uu) = \frac{1}{2}\left(\nabla \uu + (\nabla\uu)^T\right)$ is the strain-rate tensor. 
Bold-face characters denote vectors or tensors, and $\mathbf{I}$ is the identity tensor in three dimensions.
Furthermore, $\rho=1000 \ \mathrm{kg/m^3}$ is the CSF density
and $\mu=6.97\times 10^{-4}$ Pa is the dynamic viscosity of the CSF.
We introduce the tangential traction $\bsigpar$ as the
tangential component of the traction $\hat{\bsig}=\bsig\nn$:
\begin{equation*}
    \bsigpar = P_{\nn}(\hat{\bsig}) = (\mathbf{I} - \nn\otimes\nn)\hat{\bsig},
\end{equation*}
where $\nn$ is the outer unit normal of the boundary surface $\Gamma = \partial\Omega$,
and $P_{\nn}(\rr)$ is the tangential projection of a vector $\rr$ onto $\Gamma$.
We consider three types of boundary conditions:
(i) \emph{tangential traction} (no flow normal to the boundary $\uu \cdot \nn = 0$ with a force $\btau$ applied tangentially on the boundary $\bsigpar = \btau$),
(ii) \emph{normal pressure} (a pressure $\Tilde{p}$ applied in the normal direction: $(\mu\nabla\uu - p\mathbf{I})\nn = \Tilde{p}(t)\nn$),
and (iii) \emph{free slip} (no flow normal to the boundary $\uu \cdot \nn = 0$ and no tangential forces $\bsigpar = {\bf 0}$).
Unless noted otherwise, the free-slip condition is applied on the boundary. We assume that the system starts at rest: $\uu(\xx, t=0) = \mathbf{0}$.

\subsection*{Cilia-driven flow}
On the ciliated regions of the ventricle walls $\Gc$ (\Cref{fig:figure1}d),
we impose a constant-in-time tangential traction $\bsigpar=\btau$ to represent the net forces of the cilia acting on the CSF.
We set $\btau(\xx, \rr) = \tau \lambda(\xx) P_{\nn}(\rr)$, where the function $\lambda(\xx)$
and the sign of the vector $\rr=\pm(1, 0, 1)$ varies depending on the cilia region (see~\Cref{subsec:appendixA1}).
Both $\rr$ and the value $\tau=6.5\times 10^{-4}$ Pa were chosen by heuristic
model calibration using experimental data~\cite{Olstad2019CiliaryDevelopment}.

\subsection*{Flow induced by cardiac pulsations}\label{subsec:pressure_bcs}
To drive a cardiac-induced pulsatile flow inside the brain ventricles, we set $\Tilde{p}(t)=0$ and
$\Tilde{p}(t)=-A\sin{\omega t}$ on anterior and posterior parts, respectively, of the boundary $\Gp$~(\Cref{fig:figure1}d, blue markers).
The amplitude $A=1.5\times 10^{-3}$ Pa determines the magnitude of the pulsatile flow,
and was calibrated based on experimental data~\cite{Olstad2019CiliaryDevelopment}.
The angular cardiac frequency $\omega=2\pi f=6.97$ rad/s was set based on a measured cardiac frequency of $f=2.22$ Hz.

\subsection*{Computational model of solute transport after photoconversion}
We simulate transport of photoconverted proteins within the brain ventricles by
modeling transport of a solute with concentration $c(\xx, t)$ for $\xx\in\Omega$ and
time $t > 0$ by the advection-diffusion equation
\begin{equation}
    \pdifft{c} + \nabla\cdot\JJ = 0 \quad \text{in } \Omega,
    \label{eq:adv_diff_strong}
\end{equation}
where the total concentration flux $\mathbf{J}$ is assumed to consist of an advective and a diffusive flux:
\begin{equation*}
    \JJ = c\uu - D\nabla c .
\end{equation*}
Here, $\uu$ is the CSF velocity field, governed by the Stokes equations~\eqref{eq:stokes_eqs},
and $D$ is the molecular diffusion coefficient of the solute.
Initially, we set $c(\xx, 0)=0$ for all $\xx\in\Omega$.
The ventricular walls are assumed to be impermeable (admitting \emph{no flux}: $\JJ\cdot\nn=0$),
except at the anterior/posterior boundary $\Gamma_p$.
At the anterior/posterior boundary, we permit the solute concentration to be advected into
and out of the domain in a manner that ensures conservation of mass (see~\Cref{subsec:adv_diff_eq_weak_form} for details).

To computationally represent protein photoconversion in a region $\Omega_{\mathrm{pc}} \subset \Omega$ (corresponding to ROI 1), we prescribe a given time-dependent value to the solute concentration in this region. 
Mimicking the experimental protocol, we consider the dorsal part of the diencephalic ventricle as the subdomain $\Omega_{\mathrm{pc}}$, and set
\begin{equation}
  c(\xx, t)=\frac{\log{(1+t/a)}}{\log{(1+T/a)}}
  \qquad \xx \in \Omega_{\mathrm{pc}}, \quad \fixme{t > 0}.
\label{eq:photoconversion_curve}
\end{equation}
The value $a=65$ was chosen to fit the fluorescence intensity curve observed in the physical photoconversion experiments.

@@

\subsection*{Physical parameters}
The CSF density $\rho$ and dynamic viscosity $\mu$ were set following Bloomfield \textit{et al.}~\cite{Bloomfield1998EffectsFluid}~(\Cref{tab:parameter_table}). We considered large, medium and small-sized molecules by three different diffusion coefficients $D_1, D_2$ and $D_3$ (\Cref{tab:parameter_table}). Experimental data~\cite{Olstad2019CiliaryDevelopment} indicate \fixme{velocities} proximate to the ventricular walls of $27.4 \pm 5.4$ \textmu m/s and $2.6 \pm 0.6$ \textmu m/s in the dorsal and ventral regions of the diencephalic ventricle, respectively. In the telencephalic ventricle, the velocities were estimated at $4.7 \pm 1.3$ \textmu m/s. We calibrated $\tau$ and $A$ heuristically by comparing these experimental velocity values with the simulated, maximum velocity magnitude. In addition, particle tracking data of pulsatile motion was used in the calibration process to determine the relative contributions of the cardiac pulsations, as compared to the cilia-driven flow, to the velocity magnitude.
\mer{MER: What velocities? Net velocities? Velocity magnitude aka speed? Also, haven't you said this part about the experimental data already? Either place the details there, or all here. I suggest there.}
\mer{MER: I suggest you introduce parameter values when introduced, and move the parts relevant for cilia/cardiac flow to where you mentioned it first, and then merge this section with the Estimation of diffusion .... Then remove this section and \Cref{tab:parameter_table}.}
\begin{table}[h!]
    \centering
    \begin{tabular}{lcccc}
    \toprule
        Parameter & Symbol & Value & Unit & Reference \\
        \midrule
        Fluid density & $\rho$ & $1000$ & $\mathrm{kg/m^3}$ & \cite{Bloomfield1998EffectsFluid} \\
        Dynamic viscosity & $\mu$ & $\num{6.97e-04}$ & $\mathrm{Pa\cdot s}$ & \cite{Bloomfield1998EffectsFluid} \\
        Cardiac frequency & $f$ & 2.22 & Hz &  \cite{Olstad2019CiliaryDevelopment}\\
        Angular cardiac frequency & $\omega = 2\pi f$ & 6.97 & rad/s & \cite{Olstad2019CiliaryDevelopment}\\
        Cilia tangential traction & $\tau$ & \num{6.5e-04} & Pa & * \\
        Pressure amplitude & $A$ & \num{1.5e-03} & Pa & * \\
        Exosome diffusion coefficient & $D_1$ & $\num{1.63e-12}$ & $\mathrm{m^2/s}$ & \textdagger\\
        STM+GFP diffusion coefficient  & $D_2$ & $\num{5.75e-11}$ & $\mathrm{m^2/s}$ & \textdagger\\
        Dendra2 diffusion coefficient  & $D_3$ & $\num{1.15e-10}$ & $\mathrm{m^2/s}$ & \textdagger\\
    \bottomrule
    \end{tabular}
    \caption{Physical parameters of the flow and transport models.
    (*) Calibrated based on experimental data~\cite{Olstad2019CiliaryDevelopment}.
    (\textdagger) Estimated based on experimental
    data~\cite{GuraSadovsky2017MeasurementExpansion, Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells}.}
    \label{tab:parameter_table}
\end{table}

@@

\mer{MER: @Halvor I suggest you move these two paragraphs into their respective Results sections (Results 2 and 5). Putting the bullet points above at the end of the Stokes equations is a great idea!}
When simulating transport of solute in the ventricles, we consider the same six different regions of interest that were considered in the photoconversion experiments. In region number one (the photoconversion region), we use the function~\eqref{eq:photoconversion_curve} to enforce the concentration such that it represents the experimental data of photoconversion of proteins. We consider the three different diffusion coefficients introduced in previous sections to simulate and compare results for transport of molecules of varying sizes.

@@

\subsection*{Quantities of interest}
We calculate the mean solute concentrations
\begin{equation}
    \cbar_i(t) = \frac{1}{|\Omega_i|}\int_{\Omega_i} c(\xx, t)\,\dx,
    \label{eq:c_mean_i}
\end{equation}
in the six regions of interest (ROIs) $\Omega_i$ ($i = 1, \dots, 6$) as functions of time, where $|\Omega_i|$ is the volume of the domain $\Omega_i$. The dynamics of ROI 1 follows directly from~\eqref{eq:photoconversion_curve}. In addition, we report time-to-threshold: the time $\hat{t}_i$ when the mean concentration $\cbar_i$ first exceed a threshold value $\hat{c}_i$. We use $\hat{c}_i=0.25$ for ROI 1--4 and $\hat{c}_i=0.10$ for ROI 5--6.

@@

Transport of solute is quantified by the profile of the mean concentrations $\cbar_i(t)$ in the regions of interest (ROIs) $\Omega_i$ as functions of time~(\Cref{fig:fig2}b--d).
An illustration of the ROIs is provided in \Cref{fig:fig2}e. The dynamics of ROI 1 follow from imposing~\Cref{eq:photoconversion_curve}.
For ROIs 2--6, we observe more rapid concentration increases when the diffusion coefficient increases. 
For all three diffusion coefficients, the dynamics of ROI 6 is the slowest.
Region 5 exhibits a more rapid increase in mean concentration, and an even faster increase is observed for ROIs 2--4.
The profiles of the latter three regions are similar, displaying only small differences in all of the three plots.
\lyng{HH: Add some quantification in terms of percentages, e.g. final values of $\cbar_i$ for the different $D_i$?}
In region 6, the threshold is never exceeded for $D_1$, with the mean concentration close to zero at the end of the simulation. 
To quantify the balance of advection versus diffusion as transport mehcanisms on a global scale,
we calculate the Péclet numbers $\mathrm{Pe}_1=519$, $\mathrm{Pe}_2=14.7$ and $\mathrm{Pe}_3=7.4$ for $D_1$, $D_2$ and $D_3$, respectively.
Since $\mathrm{Pe}>1$, this indicates that advection is the dominant transport mechanism for the molecules considered.

@@

\subsection*{Loss of cilia motility in specific subpopulation slows down transport}
We previously identified three distinct ciliated cell lineages in the
zebrafish larval brain~\cite{Olstad2019CiliaryDevelopment, DGama2025MotileBrain}.
To identify the role of these lineages on CSF transport, we modified flow boundary conditions
in the three different ciliated regions (\Cref{fig:figure5_compare_cilia_modifications}a) and simulated
CSF flow using the Stokes equations. We identified that the velocities were most strongly
affected when removing cilia in the dorsal diencephalic cilia,
as the maximum velocity dropped from 27.7 $\mu$m/s to 8.0 $\mu$m/s.
In contrast, removing telencephalic and ventral diencephalic had minimal
impact and resulted in maximum velocities of 27.7 $\mu$m/s or 27.2 $\mu$m/s, respectively.

@@

\subsection*{The roles of advection and diffusion in intraventricular solute transport}
To investigate the roles of advection and diffusion in solute transport, we quantified transport
based on changes in fluorescence intensity $\Delta F$ (experiments) and
mean concentration $\cbar$ (simulations) as functions of time.
Even though we identified very similar dynamics in simulations and
\emph{in vivo} experiments, the dynamics of the simulations were generally
more rapid and resulted in higher total amounts.
Differences between experiments and simulations may relate to
ventricular size, as we showed that modifying the duct volumes impacts
solute distribution and the geometry used for the model was larger than in experiments.
Alternatively, it may relate to the way we defined the photoconversion site (region of interest 1)
in the simulations, since the volume of the photoconversion region determines
the total amount of solute, and thereby the rate of transport. Another source
of discrepancy may relate to the permeability of the ventricular wall,
which we have ignored in our simulation. Reports have shown that the
neuroepithelium lining the walls of brain ventricles in 24h old zebrafish
is permeable to dyes smaller than 70 kDa~\cite{Chang2012AnNeuroepithelium}.
Since Dendra2 is around 26 kDa in size~\cite{Gurskaya2006EngineeringLight},
and thus under this reported limit, one could assume that if Dendra2
leaks out of the brain ventricles, the dynamics become slower.

@@
